---
title: "How does vaccine distribution relate to GDP?"
output:
  pdf_document: default
  html_document: default
---

The data in this file comes from ["Our World in Data"](https://ourworldindata.org/grapher/covid-vaccinations-vs-gdp-per-capita), and is updated daily.  



```{r echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
covid_vaccines_orig <- read_csv("covid-vaccinations-vs-gdp-per-capita.csv")

covid_vaccines <- covid_vaccines_orig %>%
  select(country = Entity,
         date = Day,
         vaccines_per_hundred = total_vaccinations_per_hundred,
         gdp_per_capita = `GDP per capita, PPP (constant 2011 international $)`,
         continent = Continent) 

```


```{r echo = FALSE, eval = FALSE}
#NOT RUN
# What is Year and Year_1?
head(covid_vaccines)

# dimension of data
dim(covid_vaccines)

# Explore missing values 
covid_vaccines %>% 
  map_dbl(~sum(is.na(.)))
# say something about the countries with missing values

# note that there are varying numbers of rows for each country appearing in the data
covid_vaccines %>%
  count(country)

# the final date is different for each country too
covid_vaccines %>%
  group_by(country) %>%
  summarize(final_date = max(date))

```

Oh no. This data is a little bit messy! There are inconsistent numbers of rows for each country, there are missing values, and each country's final date is different.

## Examining the relationship between vaccines and GDP

Let's create some plots just for the highest reported vaccine dose count for each country.

```{r echo = FALSE}
# There are a few ways that we could get the "final" "vaccination per hundreds" number for each country.
# 1. We could filter to the last date for each country
# 2. We could filter to the highest vaccines_per_hundred number for each country
# let's do the latter:
total_covid_vaccines <- covid_vaccines %>%
  group_by(country) %>%
  filter(vaccines_per_hundred == max(vaccines_per_hundred)) %>%
  ungroup()
```

Here are some scatterplots. I encourage you to play with this yourself and make some other plots that might be interesting (e.g., the vaccines over time for different countries!)

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.align="center", out.height = "70%", out.width = "70%"}

#What about a log scale? How do  you interpret a log scale version of this plot?
total_covid_vaccines %>%
  ggplot() +
  geom_point(aes(x = gdp_per_capita / 1000, 
                 y = vaccines_per_hundred, color = continent), 
             alpha = 0.9) +
  # Annotate a few countries
  geom_text(aes(x = gdp_per_capita / 1000, 
                y = vaccines_per_hundred, label = country), 
            check_overlap = TRUE, 
            col = "grey50",
            hjust = 0, 
            alpha = 0.5) +
  scale_x_log10("GDP per capita ($1000)") +
  scale_y_continuous("Vaccines per hundred people") +
  scale_color_brewer(palette = 4, type = "qual") +
  # Annotate a few countries
  theme_classic()
```


